---
title: "Intro-Spatial"
author: "Joshua Reyling"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("setup.R")
```

```{r}
counties <- counties(state = 'CO')

roads <- roads(state = 'CO', county = 'Larimer')
```

```{r}
tmap_mode("view")
```

```{r}
pdrHwy <- roads %>% 
  filter(FULLNAME == 'Poudre Canyon Hwy')
```

```{r}
pdrPnts <- data.frame(name = c("Mishawaka", "Rustic", "Blue Lake Trailhead"),
                            long = c(-105.35634, -105.58159, -105.85563),
                            lat = c(40.68752, 40.69687, 40.57960))

pdrPnts_sf <- st_as_sf(pdrPnts, coords = c("long", "lat"), crs = 4326)
pdrPntsProject <- st_transform(pdrPnts_sf, st_crs(counties))

pdrPntsProject <- mutate(pdrPntsProject, ID = c(1:3))
```

```{r}
write_sf(pdrHwy, 'data/poudreHwy.shp')

write_sf(pdrPntsProject, 'data/poudrePoints.shp')
```

## Exercises

#### 1. Filter Out the counties data set to only include Larimer, Denver, and Pueblo counties

```{r}
countiesFiltered <- counties %>% 
  filter(NAME %in% c('Larimer', 'Denver', 'Pueblo'))
```

#### 2. Make a map of the `counties` data colored by county area. Make a second map of counties colored by their total area of water.

```{r}
# colored by land area
tm_shape(countiesFiltered) +
  tm_polygons(col = 'ALAND', )

tm_shape(countiesFiltered) +
  tm_polygons(col = 'AWATER')
```

#### 4. Make a barplot comparing the elevation of your 3 points in the Poudre Canyon (note: explore the `extract()` function in the `terra` package).

```{r}
elevation <- get_elev_raster(counties, z = 7)
elevation <- rast(elevation)
names(elevation) <- 'Elevation'

elevationCrop <- crop(elevation, ext(pdrHwy))

writeRaster(elevationCrop, 'data/poudreElevation.tif', overwrite=TRUE)

pdrPointElevation <- extract(x= elevationCrop, 
                             y= pdrPntsProject,
                             xy= TRUE
                             ) 

pdrPointElevation %>%
  left_join(pdrPntsProject, by= "ID") %>% 
  ggplot() +
    geom_col(mapping = aes(x=name, y=Elevation))
```

#### 4. Why are there 4 features in our Poudre Canyon Highway variable instead of 1?

The highway line is broken into segments, each having its own entry in the data set.
