---
title: "Intro-Spatial"
author: "Joshua Reyling"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("setup.R")
```

##### Create County object

```{r}
# Define object for counties in Colorado
counties <- counties(state = 'CO')
```

##### Create Poudre Highway line object

```{r}
# Create a line object of roads in Larimer county
roads <- roads(state = 'CO', county = 'Larimer')

# subset roads to only Poudre Highway
pdrHwy <- roads %>% 
  filter(FULLNAME == 'Poudre Canyon Hwy')
```

##### Create Points along Poudre Highway

```{r}
# create a dataframe
pdrPnts <- data.frame(name = c("Mishawaka", "Rustic", "Blue Lake Trailhead"),
                            long = c(-105.35634, -105.58159, -105.85563),
                            lat = c(40.68752, 40.69687, 40.57960))

# create a spatial object from the dataframe
pdrPnts_sf <- st_as_sf(pdrPnts, coords = c("long", "lat"), crs = 4326)

# Project created points
pdrPntsProject <- st_transform(pdrPnts_sf, st_crs(counties))

# add ID field to points (because we didn't do it before)
pdrPntsProject <- mutate(pdrPntsProject, ID = c(1:nrow(pdrPnts)))
```

##### Save highway line and points as shapefiles

```{r}
write_sf(pdrHwy, 'data/poudreHwy.shp')

write_sf(pdrPntsProject, 'data/poudrePoints.shp')
```

## Exercises

#### 1. Filter Out the counties data set to only include Larimer, Denver, and Pueblo counties

```{r}
countiesFiltered <- counties %>% 
  filter(NAME %in% c('Larimer', 'Denver', 'Pueblo'))
```

#### 2. Make a map of the `counties` data colored by county area. Make a second map of counties colored by their total area of water.

##### *colored by land area*

```{r}
tm_shape(countiesFiltered) +
  tm_polygons(col = 'ALAND', )
```

##### *colored by water area*

```{r}
tm_shape(countiesFiltered) +
  tm_polygons(col = 'AWATER')
```

#### 4. Make a barplot comparing the elevation of your 3 points in the Poudre Canyon (note: explore the `extract()` function in the `terra` package).

```{r}
elevation <- get_elev_raster(counties, z = 7)
elevation <- rast(elevation)
names(elevation) <- 'Elevation'

elevationCrop <- crop(elevation, ext(pdrHwy))

writeRaster(elevationCrop, 'data/poudreElevation.tif', overwrite=TRUE)

pdrPointElevation <- extract(x= elevationCrop, 
                             y= pdrPntsProject,
                             xy= TRUE
                             ) 

?extract
```

##### *create the plot*

```{r}
pdrPointElevation %>%
  left_join(pdrPntsProject, by= "ID") %>% 
  ggplot() +
    geom_col(mapping = aes(x=name, y=Elevation, fill=name))
```

#### 4. Why are there 4 features in our Poudre Canyon Highway variable instead of 1?

The highway has three small offshoots with the same name, each having its own entry in the data set. You can see this by plotting this line with color based on the LINEARID field:

```{r}
tm_shape(pdrHwy) +
  tm_lines(col='LINEARID', palette='Set2')
```
